---
title: "Topic 02 code-along: Signal versus noise in an MND trial"
output: html_document
---

This is a fictional MND trial inspired by work on neurodegenerative clinical trials. Higher change scores are better. It is not an MND-SMART result.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
mnd <- read_csv("data/topic02_synthetic.csv", show_col_types = FALSE)
```

## 1. Inspect completeness and distributions

```{r inspect}
mnd |> count(trial_arm, completed_followup)

# TODO: plot change_in_function_score by trial_arm.
```

## 2. Estimate the difference

```{r estimate}
mnd |>
  group_by(trial_arm) |>
  summarise(
    participants = sum(!is.na(change_in_function_score))
    # TODO: calculate mean change and SD.
  )
```

## 3. Confidence interval and test

```{r inference}
# TODO: run t.test(change_in_function_score ~ trial_arm, data = mnd)
# and use broom::tidy() to obtain an estimate and confidence interval.
```

## 4. Samples vary

```{r resampling}
# TODO: repeatedly sample 60 completed participants and calculate
# the Active minus Placebo difference in mean change.
```

## 5. Interpretation

Report the estimated effect and confidence interval before the p-value. Does statistical uncertainty settle whether the effect is clinically worthwhile?
