---
title: "Topic 02 model solutions: Signal versus noise in an MND trial"
output: html_document
---

This fictional trial contains no real participants and is not an MND-SMART result. Higher change scores are better.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
mnd <- read_csv("data/topic02_synthetic.csv", show_col_types = FALSE)
completed <- mnd |> filter(!is.na(change_in_function_score))
```

## Inspect and estimate

```{r inspect}
mnd |> count(trial_arm, completed_followup)

completed |>
  ggplot(aes(trial_arm, change_in_function_score)) +
  geom_boxplot() +
  labs(x = NULL, y = "Change in fictional function score")

completed |>
  group_by(trial_arm) |>
  summarise(n = n(), mean_change = mean(change_in_function_score), sd_change = sd(change_in_function_score))
```

## Confidence interval and p-value

```{r inference}
trial_test <- t.test(change_in_function_score ~ trial_arm, data = completed)
tidy(trial_test, conf.int = TRUE)
```

Because R orders the arms alphabetically here, the estimate is Active minus Placebo. A positive difference means less average decline with the active treatment.

## Repeated samples

```{r resampling}
set.seed(202)
sample_effect <- function(data, size = 60) {
  s <- data |> slice_sample(n = size)
  means <- tapply(s$change_in_function_score, s$trial_arm, mean)
  unname(means["Active"] - means["Placebo"])
}

sampling_results <- tibble(estimate = replicate(500, sample_effect(completed)))

sampling_results |>
  ggplot(aes(estimate)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = 0, linetype = 2) +
  labs(x = "Active minus placebo difference", y = "Repeated samples")

sampling_results |> summarise(mean = mean(estimate), sd = sd(estimate), proportion_below_zero = mean(estimate < 0))
```

The estimate and interval describe effect size and precision. The p-value does not determine whether the magnitude is worthwhile, nor whether incomplete follow-up is unbiased.

