---
title: "Topic 06 code-along: Regression and ME biomarkers"
output: html_document
---

The biomarker in this dataset is fictional. The analysis is inspired by ME biomarker research but is not a diagnostic study or a DecodeME result.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
me <- read_csv("data/topic06_synthetic.csv", show_col_types = FALSE)
```

## 1. Inspect before modelling

```{r inspect}
# TODO: compare biomarker distributions by me_status.
# TODO: compare daily steps by me_status.
```

## 2. Crude association

```{r crude}
# TODO: fit biomarker_concentration_au ~ me_status.
# Interpret the ME coefficient in arbitrary units.
```

## 3. Sequential adjustment

```{r adjustment}
# TODO: add age and sex.
# TODO: then add BMI and medication.
# TODO: finally add daily_steps.
```

## 4. Compare coefficients

```{r compare}
# TODO: use broom::tidy(..., conf.int = TRUE) to compare the ME coefficient.
```

## Interpretation

Does adjustment for activity answer the same question as the crude model? Could activity be a consequence rather than a cause of ME status?

