---
title: "Topic 03 model solutions: Comparing islet-transplant groups"
output: html_document
---

The graft-function score and all records are fictional; this is not the BETA-2 score.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
islet <- read_csv("data/topic03_synthetic.csv", show_col_types = FALSE)
```

## Describe the groups

```{r describe}
islet |>
  ggplot(aes(transplant_strategy, graft_function_score)) +
  geom_boxplot() +
  labs(x = NULL, y = "Fictional graft-function score") +
  theme(axis.text.x = element_text(angle = 15, hjust = 1))

islet |>
  group_by(transplant_strategy) |>
  summarise(
    recipients = n(),
    mean_score = mean(graft_function_score),
    sd_score = sd(graft_function_score),
    insulin_independent = mean(insulin_independent_at_12_months == "Yes")
  )
```

## Two-group comparison

```{r two-groups}
two_groups <- islet |> filter(transplant_strategy != "Repeat transplant")
tidy(t.test(graft_function_score ~ transplant_strategy, data = two_groups), conf.int = TRUE)
```

## Three-group comparison

```{r anova}
fit_anova <- aov(graft_function_score ~ transplant_strategy, data = islet)
tidy(fit_anova)
TukeyHSD(fit_anova)
```

The ANOVA tests whether all three population means are equal. Tukey-adjusted comparisons indicate which pairs differ while controlling family-wise error.

## Categorical outcome

```{r categorical}
independence_table <- table(islet$transplant_strategy, islet$insulin_independent_at_12_months)
independence_table
prop.table(independence_table, margin = 1)
chisq.test(independence_table)
```

Statistical evidence should be accompanied by group-specific estimates, uncertainty, units and a clinical judgement about whether the fictional score difference matters.

