---
title: "Topic 03 model solutions: Comparing islet-transplant groups"
output: html_document
---

These guided examples use different questions from Lab 03 so students can transfer the workflow independently. All records and results are fictional.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
islet <- read_csv("data/topic03_synthetic.csv", show_col_types = FALSE)
```

## 1. Inspect and prepare

```{r inspect}
glimpse(islet)
islet |> count(transplant_strategy)
```

```{r prepare}
code_along_data <- islet |>
  select(
    recipient_id, transplant_strategy, age_years,
    hba1c_mmol_mol, daily_insulin_units,
    insulin_independent_at_12_months
  )

code_along_data |>
  summarise(across(everything(), ~ sum(is.na(.x))))
```

There are no missing values in these variables. Checking still matters because missing observations can reduce precision or bias a comparison.

## 2. Guided example: two means

```{r two-group-explore}
code_along_data |>
  ggplot(aes(insulin_independent_at_12_months, hba1c_mmol_mol)) +
  geom_boxplot() +
  labs(x = "Insulin independent at 12 months", y = "HbA1c (mmol/mol)")

hba1c_summaries <- code_along_data |>
  group_by(insulin_independent_at_12_months) |>
  summarise(
    recipients = n(),
    mean_hba1c = mean(hba1c_mmol_mol),
    sd_hba1c = sd(hba1c_mmol_mol),
    .groups = "drop"
  )
hba1c_summaries
```

```{r two-group-test}
hba1c_test <- t.test(
  hba1c_mmol_mol ~ insulin_independent_at_12_months,
  data = code_along_data
)
hba1c_result <- tidy(hba1c_test, conf.int = TRUE)
hba1c_result
```

R orders the groups alphabetically, so the estimate is **No minus Yes**. Interpret it with the group means and interval. The p-value does not measure clinical importance or prove causation.

## 3. Guided example: three means

```{r three-group-explore}
code_along_data |>
  ggplot(aes(transplant_strategy, daily_insulin_units)) +
  geom_boxplot() +
  labs(x = NULL, y = "Daily insulin use (units/day)") +
  theme(axis.text.x = element_text(angle = 15, hjust = 1))

insulin_summaries <- code_along_data |>
  group_by(transplant_strategy) |>
  summarise(
    recipients = n(),
    mean_daily_units = mean(daily_insulin_units),
    sd_daily_units = sd(daily_insulin_units),
    .groups = "drop"
  )
insulin_summaries
```

```{r anova}
insulin_anova <- aov(
  daily_insulin_units ~ transplant_strategy,
  data = code_along_data
)
tidy(insulin_anova)
```

The overall ANOVA tests whether all three population means are equal; it does not identify the differing pairs.

```{r tukey}
insulin_tukey <- TukeyHSD(insulin_anova)
insulin_tukey
```

`diff` is the named mean difference, `lwr` and `upr` form the simultaneous interval, and `p adj` is multiplicity adjusted. Tukey adjustment controls the family-wise error rate.

## 4. Guided example: categories

```{r categorical-prepare}
categorical_data <- code_along_data |>
  mutate(
    age_group = if_else(age_years < 45, "Under 45 years", "45 years or older"),
    age_group = factor(age_group, levels = c("Under 45 years", "45 years or older"))
  )
```

```{r categorical-explore}
age_independence_table <- table(
  categorical_data$age_group,
  categorical_data$insulin_independent_at_12_months
)
age_independence_table
prop.table(age_independence_table, margin = 1)

categorical_data |>
  ggplot(aes(age_group, fill = insulin_independent_at_12_months)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL, y = "Proportion of recipients",
    fill = "Insulin independent\nat 12 months"
  )
```

```{r categorical-test}
age_independence_test <- chisq.test(age_independence_table)
age_independence_test
age_independence_test$expected
```

Check expected counts before relying on the approximation. Row proportions describe the pattern; the p-value tests independence. This observational association cannot establish causation.

## 5. Connection to Lab 03

The Lab independently compares graft-function scores between transplant strategies, adds assumption checks including Levene's test, and examines transplant strategy against insulin independence. The Code Along demonstrates the methods without supplying those analyses or conclusions.

The Lab's `graft_function_score` is fictional and is not the BETA-2 score.
