---
title: "Topic 03 code-along: Comparing islet-transplant groups"
output: html_document
---

This guided Code Along uses a synthetic dataset inspired by islet-transplant research. All records and results are fictional.

The examples deliberately use **different research questions from Lab 03**. Learn the workflow here, then apply it independently to new comparisons in the Lab.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
islet <- read_csv("data/topic03_synthetic.csv", show_col_types = FALSE)
```

## The comparison workflow

For each example: **define** the question and variables, **prepare** the data, **explore** the groups, **compare** them, then **interpret** magnitude, uncertainty, and evidence.

## 1. Inspect and prepare

```{r inspect}
# TODO: inspect islet with glimpse().
# TODO: count recipients by transplant_strategy.
```

Create `code_along_data` containing `recipient_id`, `transplant_strategy`, `age_years`, `hba1c_mmol_mol`, `daily_insulin_units`, and `insulin_independent_at_12_months`. Then count missing values in every selected column.

```{r prepare}
# TODO: create code_along_data with select().
# TODO: count missing values with summarise() and across().
```

## 2. Guided example: two means

**Question:** Does mean HbA1c differ between recipients who are and are not insulin independent at 12 months?

- Outcome: `hba1c_mmol_mol` (continuous)
- Group: `insulin_independent_at_12_months` (two independent groups)

Create a boxplot and calculate each group's number of recipients, mean HbA1c, and standard deviation.

```{r two-group-explore}
# TODO: boxplot hba1c_mmol_mol by insulin-independence status.
# TODO: calculate n, mean HbA1c, and SD by group.
```

Run a Welch independent-samples t-test and use `broom::tidy()` with `conf.int = TRUE`.

```{r two-group-test}
# TODO: run t.test(hba1c_mmol_mol ~
#   insulin_independent_at_12_months, data = code_along_data).
# TODO: tidy the result with a confidence interval.
```

Report group means and the estimated difference before the p-value. What does the interval say about precision? Why does this not establish clinical importance or causation?

## 3. Guided example: three means

**Question:** Does mean daily insulin use differ across the three transplant strategies?

- Outcome: `daily_insulin_units` (continuous)
- Group: `transplant_strategy` (three independent groups)

```{r three-group-explore}
# TODO: boxplot daily_insulin_units by transplant_strategy.
# TODO: calculate n, mean, and SD for each strategy.
```

Fit a one-way ANOVA and inspect the overall test.

```{r anova}
# TODO: fit daily_insulin_units ~ transplant_strategy with aov().
# Save it as insulin_anova, then inspect it with tidy().
```

The overall test does not identify which pairs differ. Use Tukey-adjusted comparisons.

```{r tukey}
# TODO: run TukeyHSD(insulin_anova).
```

Interpret the pairwise estimates, simultaneous intervals, and adjusted p-values. Why are adjusted results preferable to several unadjusted t-tests?

## 4. Guided example: categories

**Question:** In this synthetic dataset, is age group associated with insulin independence at 12 months?

For teaching purposes, create **Under 45 years** and **45 years or older** groups. Categorising age discards information, so this derived variable is used only to demonstrate the workflow.

```{r categorical-prepare}
# TODO: use mutate() and if_else() to create age_group.
# Store the result as categorical_data.
```

Construct a contingency table, calculate row proportions, and draw a proportional bar chart.

```{r categorical-explore}
# TODO: create age_independence_table with table().
# TODO: display counts and row proportions.
# TODO: draw a bar chart using position = "fill".
```

Run a chi-squared test and inspect expected cell counts.

```{r categorical-test}
# TODO: run chisq.test(age_independence_table) and save it.
# TODO: print the result and inspect $expected.
```

Are expected counts adequate? Describe the proportions as well as the p-value. Why can association not establish that age causes insulin independence?

## 5. Prepare for Lab 03

The Lab transfers this workflow to **different questions** using the same data:

- compare fictional graft-function scores between transplant-strategy groups;
- justify assumption checks, including Levene's test;
- examine transplant strategy and insulin independence;
- choose the relevant visualisations;
- distinguish effect size, statistical significance, and clinical significance.

Keep this script as a method reference, but do not reuse its conclusions as Lab answers. `graft_function_score` is an invented teaching measure, not the validated BETA-2 score.
