---
title: "Topic 05 code-along: Measuring developmental risk"
output: html_document
---

This fictional longitudinal dataset is inspired by CHILDS. It does not show that a prenatal exposure causes a developmental outcome.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
theme_set(theme_minimal())
childs <- read_csv("data/topic05_synthetic.csv", show_col_types = FALSE)
```

## 1. Define the analysis population

```{r cohort}
childs |> count(observed_at_followup)

observed <- childs |>
  filter(observed_at_followup == "Yes")
```

## 2. Risks in exposed and unexposed groups

```{r risks}
risk_summary <- observed |>
  group_by(prenatal_exposure) |>
  summarise(
    children = n(),
    events = sum(developmental_support_flag == "Yes")
    # TODO: calculate risk.
  )

risk_summary
```

## 3. Absolute and relative effects

```{r effects}
# TODO: calculate the risk difference and risk ratio.
# TODO: calculate odds in each group and the odds ratio.
```

## 4. Examine framing and missing follow-up

```{r missingness}
# TODO: compare follow-up completion by deprivation quintile and exposure.
```

## Interpretation

Report both groups' risks using the same denominator, followed by absolute and relative differences. Explain why the risk ratio is not automatically causal.
