---
title: "Topic 05 model solutions: Measuring developmental risk"
output: html_document
---

This fictional longitudinal dataset is not a CHILDS result and does not establish causality.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
theme_set(theme_minimal())
childs <- read_csv("data/topic05_synthetic.csv", show_col_types = FALSE)
observed <- childs |> filter(observed_at_followup == "Yes")
```

## Absolute risks

```{r risks}
risk_summary <- observed |>
  group_by(prenatal_exposure) |>
  summarise(children = n(), events = sum(developmental_support_flag == "Yes"), risk = events / children)
risk_summary

risk_exposed <- risk_summary$risk[risk_summary$prenatal_exposure == "Recorded exposure"]
risk_unexposed <- risk_summary$risk[risk_summary$prenatal_exposure == "No recorded exposure"]
```

## Effect measures

```{r measures}
tibble(
  measure = c("Risk difference", "Risk ratio", "Odds ratio"),
  estimate = c(
    risk_exposed - risk_unexposed,
    risk_exposed / risk_unexposed,
    (risk_exposed / (1-risk_exposed)) / (risk_unexposed / (1-risk_unexposed))
  )
)
```

The risk difference communicates additional events per child observed; multiplying by 100 gives additional events per 100. The risk ratio compares probabilities. The odds ratio compares event odds and is more distant from 1 when outcomes are not rare.

## Follow-up

```{r followup}
childs |>
  group_by(prenatal_exposure, deprivation_quintile) |>
  summarise(n = n(), followup_proportion = mean(observed_at_followup == "Yes"), .groups = "drop") |>
  ggplot(aes(deprivation_quintile, followup_proportion, colour = prenatal_exposure)) +
  geom_line() + geom_point() +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Deprivation quintile", y = "Observed at follow-up", colour = "Prenatal exposure")
```

Balanced reporting gives risks in both groups, the same follow-up period, absolute and relative comparisons, and acknowledges confounding and differential follow-up.

