---
title: "Topic 07 model solutions: Predicting colorectal-cancer progression"
output: html_document
---

The records and simplified genomic categories are fictional and reproduce no real study result.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
crc <- read_csv("data/topic07_synthetic.csv", show_col_types = FALSE) |>
  mutate(progressed = progression_within_2_years == "Yes")
development <- crc |> filter(cohort == "Development")
validation <- crc |> filter(cohort == "External validation")
```

## Develop the model

```{r model}
progression_model <- glm(
  progressed ~ age_years + tumour_stage + msi_status + apc_function_group + molecular_subtype,
  data = development, family = binomial()
)
tidy(progression_model, exponentiate = TRUE, conf.int = TRUE)
```

## Predict development and external-validation cohorts

```{r predictions}
development <- development |> mutate(predicted_risk = predict(progression_model, newdata = development, type = "response"))
validation <- validation |> mutate(predicted_risk = predict(progression_model, newdata = validation, type = "response"))
```

```{r auc}
auc_rank <- function(truth, probability) {
  n1 <- sum(truth); n0 <- sum(!truth)
  (sum(rank(probability)[truth]) - n1*(n1+1)/2) / (n1*n0)
}

tibble(
  cohort = c("Development", "External validation"),
  auc = c(auc_rank(development$progressed, development$predicted_risk),
          auc_rank(validation$progressed, validation$predicted_risk)),
  brier_score = c(mean((development$progressed-development$predicted_risk)^2),
                  mean((validation$progressed-validation$predicted_risk)^2)),
  mean_prediction = c(mean(development$predicted_risk), mean(validation$predicted_risk)),
  observed_risk = c(mean(development$progressed), mean(validation$progressed))
)
```

## Calibration in the validation cohort

```{r calibration}
calibration <- validation |>
  mutate(risk_group = ntile(predicted_risk, 5)) |>
  group_by(risk_group) |>
  summarise(n = n(), mean_prediction = mean(predicted_risk), observed_risk = mean(progressed))

calibration
calibration |>
  ggplot(aes(mean_prediction, observed_risk)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_point(size = 3) + geom_line() +
  coord_equal(xlim = c(0, .8), ylim = c(0, .8)) +
  labs(x = "Mean predicted risk", y = "Observed progression proportion")
```

A coefficient estimates a conditional association; prediction asks how accurately the complete model performs for unseen patients. External validation protects this assessment from optimism created during model development.

