---
title: "Topic 07 code-along: Predicting colorectal-cancer progression"
output: html_document
---

This fictional colorectal-cancer dataset contains simplified genomic categories. It does not reproduce any real cohort, genotype or research result.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
crc <- read_csv("data/topic07_synthetic.csv", show_col_types = FALSE)
```

## 1. Preserve the validation cohort

```{r split}
development <- crc |> filter(cohort == "Development")
validation <- crc |> filter(cohort == "External validation")

development |> count(progression_within_2_years)
validation |> count(progression_within_2_years)
```

## 2. Fit on development data only

```{r model}
# TODO: fit logistic regression for progression using age, stage,
# MSI status, APC function group and molecular subtype.
```

## 3. Predict unseen patients

```{r predict}
# TODO: add predicted probabilities to development and validation.
```

## 4. Discrimination and calibration

```{r evaluate}
# TODO: compare average predicted risk with observed outcome prevalence.
# TODO: group validation predictions into risk bands and assess calibration.
```

## 5. Estimation versus prediction

Does a coefficient answer the same question as performance for a new patient? Why should the external-validation cohort remain untouched until evaluation?

