---
title: "Topic 04 code-along: Correlation and visualisation in R"
output: html_document
---

This guided Code Along uses synthetic real-world healthcare data. All records and results are fictional.

We will investigate the relationship between **age** and **baseline severity**. Lab 04 uses a different relationship from the same dataset so that you can apply the workflow independently.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())

rwe <- read_csv("data/topic04_synthetic.csv", show_col_types = FALSE)
```

## 1. Define the question

**Research question:** In this synthetic sample, is age associated with baseline severity score?

- Explanatory variable: `age_years`
- Outcome variable: `baseline_severity_score`
- Both variables are continuous.

Correlation treats the variables symmetrically, but naming their roles helps us describe the healthcare question clearly. An association would not establish that age causes severity.

## 2. Prepare the analysis dataset

Use `select()` to retain `patient_id`, `age_years`, and `baseline_severity_score`. Use `filter()` to keep complete observations, and save the result as `correlation_data`.

```{r prepare}
# TODO: inspect rwe with glimpse().

# TODO: use select() and filter() to create correlation_data.

# TODO: compare nrow(rwe) and nrow(correlation_data).
```

Were any observations removed? Why should completeness be checked even when no values appear to be missing?

## 3. Construct an informative scatterplot

Create a scatterplot with age on the x-axis and baseline severity on the y-axis. Use transparency to reduce overplotting and add clear labels.

```{r scatterplot}
# TODO: create the scatterplot with ggplot() and geom_point().
```

Describe the direction, strength, and shape of the relationship. Are there unusual observations? Does a straight-line summary appear reasonable?

## 4. Add a fitted line

Add a linear fitted line and confidence band using `geom_smooth(method = "lm")`.

```{r fitted-line}
# TODO: recreate the scatterplot and add geom_smooth(method = "lm").
```

The fitted line is a visual summary of the trend. It does not demonstrate a causal effect.

## 5. Calculate and test Pearson's correlation

Use `cor()` to calculate Pearson's correlation coefficient. Then run `cor.test()` and save the full test object as `age_severity_test`.

```{r correlation}
# TODO: calculate Pearson's r with cor().

# TODO: run cor.test(age_years, baseline_severity_score,
#                    method = "pearson") and save the result.

# TODO: print age_severity_test.
```

Interpret the sign and magnitude of (r). What do the confidence interval and p-value add? Does a small p-value imply a strong or important relationship?

## 6. Extract results from the test object

First extract values directly from the `cor.test()` object. Then use `broom::tidy()` to create a one-row results table.

```{r extract}
# TODO: extract age_severity_test$estimate,
# age_severity_test$conf.int, and age_severity_test$p.value.

# TODO: tidy age_severity_test and save it as correlation_result.
```

## 7. Add a dynamic figure caption

Create a caption using the values in `correlation_result`. Round the coefficient and interval to two decimal places and format the p-value appropriately.

```{r caption}
# TODO: create correlation_caption with sprintf() or paste0().

# TODO: add labs(caption = correlation_caption) to the scatterplot.
```

A dynamic caption updates automatically if the data or analysis changes, helping keep figures and statistical results consistent.

## 8. Interpret in context

Write a short conclusion that:

1. describes the visible pattern;
2. reports Pearson's (r), its 95% confidence interval, and p-value;
3. distinguishes the strength of evidence from the strength of association;
4. notes that the analysis is unadjusted;
5. avoids claiming that age causes baseline severity.

## 9. Prepare for Lab 04

The Lab uses the same dataset but a different association: **age and comorbidity count**. You will independently choose and justify the analysis, draw a DAG, assess whether Pearson correlation is appropriate, and compare visual, numerical, and causal interpretations.
