---
title: "Topic 04 model solutions: Association, confounding and causation"
output: html_document
---

This is fictional real-world evidence designed to demonstrate confounding by indication.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
rwe <- read_csv("data/topic04_synthetic.csv", show_col_types = FALSE) |>
  mutate(died = outcome_90_days == "Died")
```

## Crude association

```{r crude}
crude_risks <- rwe |>
  group_by(treatment_received) |>
  summarise(patients = n(), deaths = sum(died), mortality_risk = mean(died))
crude_risks
```

The treated group has higher crude mortality. That observation alone does not tell us what would have happened to the same eligible patients under usual care.

## Treatment selection

```{r selection}
rwe |>
  group_by(treatment_received) |>
  summarise(mean_severity = mean(baseline_severity_score), mean_age = mean(age_years), mean_comorbidity = mean(comorbidity_count))

rwe |>
  ggplot(aes(treatment_received, baseline_severity_score)) +
  geom_boxplot() +
  labs(x = NULL, y = "Baseline severity score")
```

## Stratified risks

```{r stratified}
rwe <- rwe |>
  mutate(severity_group = cut(baseline_severity_score, c(0, 35, 55, 100), labels = c("Lower", "Middle", "Higher")))

rwe |>
  group_by(severity_group, treatment_received) |>
  summarise(n = n(), mortality_risk = mean(died), .groups = "drop")
```

## Crude and adjusted regression

```{r models}
crude_model <- glm(died ~ treatment_received, data = rwe, family = binomial())
adjusted_model <- glm(died ~ treatment_received + baseline_severity_score + age_years + comorbidity_count + healthcare_access, data = rwe, family = binomial())

bind_rows(
  tidy(crude_model, exponentiate = TRUE, conf.int = TRUE) |> mutate(model = "Crude"),
  tidy(adjusted_model, exponentiate = TRUE, conf.int = TRUE) |> mutate(model = "Adjusted")
) |>
  filter(str_detect(term, "treatment_received")) |>
  select(model, term, estimate, conf.low, conf.high, p.value)
```

Severity affects both treatment selection and outcome, distorting the crude comparison. Even the adjusted estimate needs causal assumptions: correct time ordering and model form, adequate measurement, positivity, consistency and no important unmeasured confounding.

