---
title: "Topic 04 code-along: Association, confounding and causation"
output: html_document
---

This fictional real-world treatment dataset was designed to contain confounding by indication. It does not establish the effect of any real treatment.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
rwe <- read_csv("data/topic04_synthetic.csv", show_col_types = FALSE)
```

## 1. Draw the assumed causal structure

Before analysing, draw arrows among baseline severity, treatment and 90-day outcome. Where should age, comorbidity and healthcare access sit?

## 2. Crude association

```{r crude}
rwe |>
  count(treatment_received, outcome_90_days) |>
  group_by(treatment_received) |>
  mutate(risk = n / sum(n))

# TODO: extract and compare mortality risks.
```

## 3. Inspect treatment selection

```{r selection}
# TODO: compare baseline_severity_score across treatment groups.
```

## 4. Stratify and adjust

```{r adjusted}
rwe <- rwe |>
  mutate(severity_group = cut(baseline_severity_score,
                              breaks = c(0, 35, 55, 100),
                              labels = c("Lower", "Middle", "Higher")))

# TODO: calculate risks within severity_group.
# TODO: fit crude and adjusted logistic models.
```

## 5. Causal interpretation

Why can crude and adjusted associations point in different directions? Which assumptions are still needed for a causal claim?

