---
title: "Topic 04 model solutions: Correlation and visualisation in R"
output: html_document
---

This guided Code Along uses synthetic real-world healthcare data. All records and results are fictional. The example differs from the independent Lab 04 question.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())

rwe <- read_csv("data/topic04_synthetic.csv", show_col_types = FALSE)
```

## 1. Define and prepare

The question asks whether age is associated with baseline severity. Both variables are continuous. An observed association would not establish causation.

```{r prepare}
glimpse(rwe)

correlation_data <- rwe |>
  select(patient_id, age_years, baseline_severity_score) |>
  filter(!is.na(age_years), !is.na(baseline_severity_score))

tibble(
  stage = c("Original data", "Complete analysis data"),
  observations = c(nrow(rwe), nrow(correlation_data))
)
```

No observations are removed, but checking prevents missing values from being handled silently.

## 2. Construct the scatterplot

```{r scatterplot}
age_severity_plot <- correlation_data |>
  ggplot(aes(x = age_years, y = baseline_severity_score)) +
  geom_point(alpha = 0.35, colour = "#5b21b6") +
  labs(
    x = "Age at eligibility (years)",
    y = "Fictional baseline severity score"
  )

age_severity_plot
```

The plot suggests a weak positive, approximately linear association with considerable scatter. The large sample makes the overall pattern easier to see even though the relationship is weak.

## 3. Add a fitted line

```{r fitted-line}
age_severity_plot <- age_severity_plot +
  geom_smooth(method = "lm", se = TRUE, colour = "#d97706")

age_severity_plot
```

The line summarises the observed linear trend; it does not represent a causal effect.

## 4. Calculate and test Pearson's correlation

```{r correlation}
pearson_r <- cor(
  correlation_data$age_years,
  correlation_data$baseline_severity_score,
  method = "pearson"
)
pearson_r

age_severity_test <- cor.test(
  correlation_data$age_years,
  correlation_data$baseline_severity_score,
  method = "pearson"
)
age_severity_test
```

Pearson's (r) is positive but small. With 2,500 observations, even a weak association can produce a small p-value. Statistical evidence against zero correlation is not the same as a strong or clinically important relationship.

## 5. Extract the results

```{r extract}
age_severity_test$estimate
age_severity_test$conf.int
age_severity_test$p.value

correlation_result <- tidy(age_severity_test)
correlation_result
```

## 6. Add a dynamic caption

```{r caption}
correlation_caption <- with(
  correlation_result,
  sprintf(
    "Pearson's r = %.2f (95%% CI %.2f to %.2f), p %s",
    estimate,
    conf.low,
    conf.high,
    if_else(p.value < 0.001, "< 0.001", paste0("= ", format.pval(p.value, digits = 3)))
  )
)

final_plot <- age_severity_plot +
  labs(caption = correlation_caption) +
  theme(plot.caption = element_text(hjust = 0))

final_plot
```

Because the caption is generated from the test result, it will update if the analysis changes.

## 7. Interpretation

There is a weak positive linear association between age and the fictional baseline severity score in this synthetic sample. Pearson's coefficient, confidence interval, and p-value should be reported from the output above. The narrow interval and small p-value reflect precision from the large sample, but the coefficient remains small. The analysis is unadjusted and cannot show that greater age causes higher severity; confounding, selection, measurement, and the data-generating process remain relevant.

## 8. Connection to Lab 04

Lab 04 independently examines age and comorbidity count, adds a DAG, and asks whether Pearson correlation is suitable for a discrete count variable. These model solutions therefore demonstrate the workflow without providing the Lab result or causal interpretation.
