---
title: "Topic 08 code-along: From population evidence to decisions"
output: html_document
---

This fictional perinatal dataset represents no real screening programme or region. It is designed to explore thresholds, equity, benefits and burdens.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
population <- read_csv("data/topic08_synthetic.csv", show_col_types = FALSE)
```

## 1. Describe access and outcomes

```{r access}
# TODO: calculate screening completion and adverse-outcome risk by region
# and deprivation quintile.
```

## 2. Estimate baseline risk

```{r model}
# For teaching only, fit a risk model using age, deprivation and region.
# TODO: generate each pregnancy's predicted baseline risk.
```

## 3. Apply decision thresholds

```{r thresholds}
# TODO: for thresholds of 5%, 10%, 15% and 20%, count how many
# pregnancies would be offered support.
```

## 4. Benefits and burdens

Assume, purely for this exercise, that support reduces relative risk by 32% and that 11% of recipients experience a burden.

```{r tradeoffs}
# TODO: estimate events prevented and burdens at each threshold.
```

## 5. Decide transparently

Which threshold would you choose? State whose outcomes and preferences matter, and show how uncertainty could change your choice.

