---
title: "Topic 08 model solutions: From population evidence to decisions"
output: html_document
---

This fictional dataset represents no real perinatal screening programme, intervention or region.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
theme_set(theme_minimal())
population <- read_csv("data/topic08_synthetic.csv", show_col_types = FALSE) |>
  mutate(adverse = adverse_perinatal_outcome == "Yes")
```

## Access, outcomes and equity

```{r access}
population |>
  group_by(region) |>
  summarise(
    pregnancies = n(),
    screened = mean(screening_completed == "Yes"),
    adverse_risk = mean(adverse),
    support_received = mean(preventive_support_accepted == "Yes")
  )

population |>
  group_by(deprivation_quintile) |>
  summarise(screened = mean(screening_completed == "Yes"), adverse_risk = mean(adverse))
```

## Estimate baseline risk for the teaching exercise

The model deliberately omits post-baseline screening and treatment variables.

```{r risk-model}
risk_model <- glm(adverse ~ maternal_age_years + I((maternal_age_years-30)^2) + factor(deprivation_quintile) + region,
                  data = population, family = binomial())
population <- population |> mutate(predicted_baseline_risk = predict(risk_model, type = "response"))
```

## Threshold analysis

For illustration only, assume relative risk under support is 0.68 and the probability of a treatment burden is 0.11.

```{r thresholds}
relative_risk <- 0.68
burden_probability <- 0.11

threshold_results <- map_dfr(c(.05, .10, .15, .20), function(threshold) {
  selected <- population |> filter(predicted_baseline_risk >= threshold)
  tibble(
    threshold = threshold,
    offered = nrow(selected),
    proportion_offered = nrow(selected) / nrow(population),
    expected_events_without_support = sum(selected$predicted_baseline_risk),
    expected_events_prevented = sum(selected$predicted_baseline_risk * (1-relative_risk)),
    expected_burdens = nrow(selected) * burden_probability
  )
})
threshold_results
```

```{r threshold-plot}
threshold_results |>
  select(threshold, expected_events_prevented, expected_burdens) |>
  pivot_longer(-threshold, names_to = "outcome", values_to = "expected_count") |>
  ggplot(aes(threshold, expected_count, colour = outcome)) +
  geom_line() + geom_point() +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Decision threshold", y = "Expected count", colour = NULL)
```

No threshold is selected by evidence alone. The decision depends on how benefits, burdens, resources, missed opportunities and inequalities are valued, and whether conclusions remain stable across plausible treatment effects.

