---
title: "Topic 01 model solutions: Describing critical-care data"
output: html_document
---

This synthetic teaching dataset contains no real patient records and its results are fictional.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
theme_set(theme_minimal())
icu <- read_csv("data/topic01_synthetic.csv", show_col_types = FALSE)
```

## Variable types and completeness

`age_years` and ICU stay are continuous measurements (age has been rounded); `sex` and admission type are nominal; deprivation quintile is ordinal; and counts of comorbidities and organ supports are discrete.

```{r completeness}
glimpse(icu)
icu |> count(admission_type) |> mutate(proportion = n / sum(n))
icu |> summarise(across(everything(), ~sum(is.na(.x))))
```

## Centre, spread and shape

```{r summaries}
icu |>
  summarise(
    admissions = n(),
    missing_icu_stay = sum(is.na(icu_length_of_stay_days)),
    mean_days = mean(icu_length_of_stay_days, na.rm = TRUE),
    sd_days = sd(icu_length_of_stay_days, na.rm = TRUE),
    median_days = median(icu_length_of_stay_days, na.rm = TRUE),
    iqr_days = IQR(icu_length_of_stay_days, na.rm = TRUE)
  )
```

```{r plots, fig.width=8, fig.height=4}
icu |>
  ggplot(aes(icu_length_of_stay_days)) +
  geom_histogram(binwidth = 1, boundary = 0, colour = "white") +
  labs(x = "ICU length of stay (days)", y = "Admissions")

icu |>
  ggplot(aes(admission_type, icu_length_of_stay_days)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 25)) +
  labs(x = NULL, y = "ICU length of stay (days)")
```

The mean exceeds the median because a minority of long stays pull it upwards. Median and IQR therefore describe a typical admission more honestly, while the histogram preserves the important long right tail.

