---
title: "Topic 01 code-along: Describing critical-care data"
output: html_document
---

This code-along uses fictional ICU admissions inspired by critical-care epidemiology. The data contain no real patients and the results are not research findings.

## 1. Set up and import

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
theme_set(theme_minimal())
```

```{r import}
icu <- read_csv("data/topic01_synthetic.csv", show_col_types = FALSE)
glimpse(icu)
```

## 2. Who is represented?

```{r variable-types}
# TODO: classify age_years, sex, deprivation_quintile,
# comorbidity_count and icu_length_of_stay_days.

icu |> count(admission_type)
```

## 3. Centre and spread

```{r summaries}
icu |>
  summarise(
    admissions = n(),
    missing_icu_stay = sum(is.na(icu_length_of_stay_days))
    # TODO: add mean, SD, median and IQR of ICU stay.
  )
```

## 4. See the distribution

```{r plots}
icu |>
  ggplot(aes(x = icu_length_of_stay_days)) +
  # TODO: add a histogram with a useful bin width.
  labs(x = "ICU length of stay (days)", y = "Admissions")
```

```{r group-plot}
# TODO: compare ICU stay across admission types with a boxplot.
```

## 5. Clinical interpretation

Write two sentences reporting the typical ICU stay and its variation. What does the plot reveal that the numerical summary does not?
